Environmental stress drives intermediate plant and pollinator specialisation

### 2020-04-17 paper/manuscript.Rmd
corrections to discussion

We set out to explore whether and how environmental stress can systematically affect specialisation. 
After accounting for the pool of potential partners, we found that environmental conditions contribute to determining whether a species is a generalist or a specialist *in their community*. 
We also found that the particular effect of the environment is strongly dependent on the species.
Based on existing literature, we proposed two alternative hypotheses of how environmental stress may affect species' specialisation, and we found evidence for both of them.
Species with a large number of partners in low-stress communities were more likely to have a negative relationship and hence reduce the number of partners as stress increases. 
Contrastingly, species in our datasets with a small number of partners in low-stress communities were more likely to have a larger number of partners in more stressful communities.
In summary, environmental stress pushes species that are flexible enough to change their interaction partners towards intermediate levels of specialisation, a so-called "regression towards the mean".

Our results suggest that changes in community composition are indeed the primary channel through which the environment determines changes interaction probability. 
However, they also show that, for a large number of species, the environment may also play a substantial role in determining their level of specialisation. 
Previous research has recognised that environmental factors may help explain the changes in network structure along environmental gradients that cannot be explained by community composition [@tylianakis_habitat_2007].
However, how these two factors were linked had been elusive so far [@gravel_bringing_2018]. 
We believe that part of this difficulty could have arisen because species, and ultimately network structure, can respond in multiple, and contrasting, ways depending on the particular bioclimatic variable examined (e.g. temperature or precipitation).
Using stress to summarise the effect on species of multiple environmental gradients allowed us to detect a clear signal of the environment in species' interaction patterns.

Although both niche and neutral processes are relevant in determining species interactions, our model suggests that niche processes may be the predominant mechanism through which the environment *systematically* affects specialisation. 
First, it is unlikely that environmental stress correlates to local species abundances [@pearce_practical_2001; @sagarin_moving_2006]. 
Second, even if there is a relationship between stress and abundances, a particular environmental gradient could have a positive effect on the abundance of some species and a negative effect on others. 
Indeed, we find that within a community there is a wide range of stress values, even for the relatively limited number of species we were able to include in our analysis.

Recent research suggests that species are continuously changing their interaction partners wherever environmental conditions change in space or time [@raimundo_adaptive_2018].
So far it appears that this rewiring is primarily driven by generalist species [@ponisio_opportunistic_2017; @burkle_plant-pollinator_2013], presumably because generalist species are less sensitive to trait matching of their interaction partners [@caradonna_interaction_2017].
Our results add two important nuances to these findings. 
First, because "generalists" seem to focus on a smaller number of partners as environmental conditions deteriorate, we show that trait matching might still play a role in determining the interactions of generalist species. 
Second, and most importantly, our results suggest that only a small proportion of species are "true generalists" or "true specialists" this is, species that interact with a large or small number of partners regardless of the environmental stress, respectively. 
This pattern implies that rewiring is not exclusive of species with a large number of partners. 
Instead, at least a fraction of the species that appear to be specialist *in their communities* might be as flexible, if not more, than those with a large number of partners, effectively behaving as facultative generalists in the face of environmental change. 
These flexible "specialists" might therefore have a more significant role in network persistence than previously expected.

@@

Similarly, if the patterns we see in our models have also played a role during the evolutionary history of pollination communities, our results also help explain why only a small fraction of plant-pollinator interactions show a strong signature of deep co-evolutionary history [@hutchinson_cophylogenetic_2017].
The increases in the stress that species are predicted to experience due to rapid environmental change might further erode the co-evolutionary history of specialist species.
Communities as a whole might be in a trajectory of even more diffuse co-evolution.
For specialists, at least, the longer-term benefits of being able to interact with multiple partners might be more important than the shorter-term benefits of interacting with partners of high trait matching.

The structural implications of the "regression towards the mean" that environmental stress promotes are less clear. 
However, it is plausible to expect that nestedness, and therefore network stability, might be reduced in the face of rapid environmental change.
Determining exactly how the changes in degree caused by environmental stress reflect on systematic changes in network structure would be an interesting avenue of research. 
Answering this question would require expanding our suitability analysis to all species in the community and compare the degree distribution of networks along a gradient of stress for the community as a whole.

In conclusion, we show that the environment can affect the specialisation level of plants and pollinators in systematic ways beyond community composition. 
Species that are inflexible with their interaction partners are unlikely to persist under more stressful environmental conditions. 
However, we show that many species are flexible in regards to their specialisation levels and therefore are not inherently generalists or specialists. 
Instead, the species' level of specialisation/generalisation should be considered on a relative scale depending on where they are found and the environmental conditions at that location.

>>>>

We set out to explore whether and how environmental stress can systematically affect specialisation. 
After accounting for the pool of potential partners, we found that environmental conditions contribute to determining whether a species is a generalist or a specialist *in their community*. 
We also found that the particular effect of the environment dependends strongly on the species' inherent tendency to be a specialist or generalist.
Based on existing literature, we proposed two alternative hypotheses of how environmental stress may affect species' specialisation, and we found evidence for both.
Species with many partners in low-stress communities were more likely to have a negative relationship and hence reduce the number of partners as stress increases. 
Contrastingly, species in our datasets with few partners in low-stress communities were more likely to interact with more partners in more stressful communities.

Our results suggest that changes in community composition are indeed the primary channel through which the environment determines changes interaction probability, and this is a common mechanism through which ecological networks respond to environmental change [@tylianakis_ecological_2017]. 
However, they also show that, for a large number of species, the environment may also play a substantial role in determining their level of specialisation. 
Previous research has recognised that environmental factors may help explain the changes in network structure along environmental gradients that cannot be explained by community composition [@tylianakis_habitat_2007].
However, how these two factors were linked had been elusive so far [@gravel_bringing_2018]. 
We believe that part of this difficulty could have arisen because species, and ultimately network structure, can respond in multiple, and contrasting, ways depending on the particular bioclimatic variable examined (e.g. temperature or precipitation).
Using stress to summarise the effect on species of multiple environmental gradients allowed us to detect a clear signal of the environment in species' interaction patterns.

Recent research suggests that species are continuously changing their interaction partners wherever environmental conditions change in space or time [@raimundo_adaptive_2018].
So far it appears that this rewiring is primarily driven by generalist species [@ponisio_opportunistic_2017; @burkle_plant-pollinator_2013], presumably because generalist species are less sensitive to trait matching of their interaction partners [@caradonna_interaction_2017].
Our results add two important nuances to these findings. 
First, because "generalists" seem to focus on fewer partners as environmental conditions deteriorate, we show that trait matching might still play a role in determining the  less-flexible core interactions of generalist species. 
Second, and most importantly, our results suggest that only a small proportion of species are "true generalists" or "true specialists" this is, species that interact with a large or small number of partners regardless of the environmental stress, respectively. 
This pattern implies that rewiring is not exclusive to species with many partners. 
Instead, at least a fraction of the species that appear to be specialist *in their communities* might be as flexible, if not more, than those with a large number of partners, effectively behaving as facultative generalists in the face of environmental change. 
These flexible "specialists" might therefore have a more significant role in network persistence than previously expected.

@@

Similarly, if the patterns we see in our models have also played a role during the evolutionary history of pollination communities, our results also help to explain why only a small fraction of plant-pollinator interactions shows a strong signature of deep co-evolutionary history [@hutchinson_cophylogenetic_2017].
The increases in the stress that species are predicted to experience due to rapid environmental change might further erode the co-evolutionary history of specialist species.
Communities as a whole might be in a trajectory of even more diffuse co-evolution.
For specialists, at least, the longer-term benefits of being able to interact with multiple partners might be more important than the shorter-term benefits of interacting with partners of high trait matching.

The structural implications of the "regression towards the mean" that environmental stress promotes are less clear. 
However, it is plausible to expect that nestedness, and therefore network stability, might be reduced in the face of rapid environmental change.
Determining exactly how the changes in degree caused by environmental stress are reflected by systematic changes in network structure would be an interesting avenue of research. 
Answering this question would require expanding our suitability analysis to all species in the community and comparing the degree distribution of networks along a gradient of stress for the community as a whole.

In conclusion, we show that the environment can affect the specialisation level of plants and pollinators in systematic ways beyond community composition. 
Species that are inflexible with their interaction partners are unlikely to persist under more stressful environmental conditions. 
However, we show that many species are flexible in regards to their specialisation levels and therefore are not inherently generalists or specialists. 
Instead, the species' level of specialisation/generalisation should be considered on a relative scale depending on environmental stress at a given location.
